6CCS3PRJ \\\vspace{0.5cm} A Pragmatic Infrastructure Modelling and Verification Framework

### 2014-04-06 paper.tex
Better subsections
Fixed spellings

After taking part in the delivery of software in the highly regulated insurance industry, I noted some problems with the reliability of the software delivered. These reliability issues were generally unrelated to the application itself, but predominantly caused by the growth in complexity of the underlying application infrastructure, which was provided and managed within the company's own datacenter as a result of regulatory constraints. As the application grew in complexity over time, the number of dependent services and components needed to keep it running became unmanagable by the company's own IT staff.

@@

The tool will then execute this model as a means of system monitoring. It will verify the correctness of the live system continuously and warn of any divergence. It will also allow the attachment of metadata and metrics to nodes. This model, metadata and metrics will then serve as a useful diagnostic tool if the state of the system changes, e.g.\ a loss of connectivity between two nodes. This fine-grained and clear mechanism of communicating failure should reduce investigation and response times when issues arise, and provide the insight which prevents them from recurring.

@@

The problem of increasing software complexity is a well publicised and researched phenomenon. Software which is complex takes longer to develop, and is harder to maintain. As it affects the value proposition to the user so directly, it is often the dominant factor in determinining the feasibility of implementing a given system.

@@

Modelling of complex software has been used for a number of years as a tool of communication between developers, project managers and stakeholders. A number of tools can be used to produce clear diagrams of a system, but models provide a level of semantic meaning that allows them to be practically used to derive or verify an implementation. The Unified Modeling Language is considered the de-facto set of graphic notation techniques to create visual models of object-oriented software-intensive systems~\cite{UMLDefinition}.

@@

Studies into increased quality, maintainability and developer productivity when using MDSE techniques can be found commonly in academia but are difficult to find in industry. In a study into Industrial Adoption of Model Driven Development, it was stated that "interviewees emphasized tool immaturity, complexity and lack of usability as major barriers to adoption"~\cite{IndustryMDSE}.

@@

Nagios is in active use by over 250,000 organisations worldwide, including Amazon, AT\&T and JP Morgan Chase~\cite{NagiosUsage}. It supports a number of common montitoring architectures, including both agent based and agentless (based on the Simple Network Management Protocol) which means it can run efficiently up to a very large number of hosts~\cite{NagiosArchitectures}.

Anecdotal feedback about Nagios typically suggests that it is a mature and flexible solution for basic monitoring of large numbers of hosts, but severely lacks some higher level concepts such as integration tests of system components. It is also considered extremely time consuming and unintuitive to set up and manage, which perhaps explains its lack of adoption within smaller businesses/startups.

@@

\begin{itemize}
  \item A Technical Architect at Red Badger Consulting
  \item A System Adminstrator at HCL
\end{itemize}

@@

\paragraph{Simplicity and Pragmatism} Often the need for extensive application and infrastructure monitoring is not recognised until after the application has been deployed into a production environment. Introducing monitoring at an earlier stage, eg. in a staging environment, can highlight issues of reliability and performance before they are experienced in production. Monitoring key domain specific metrics can also aid in reasoning about the runtime behavior of the system. With these benefits in mind, the tool should be simple enough to introduce early in the development process with minimal effort on the part of the developer, it should provide pragmatic value from the start, and offer the flexibility to continue to deliver value as the system grows in complexity.

@@

When the binary is first executed, it performs node registration by making a PUT request to the monitoring tool’s REST API with the hostname of the server, and supplementary metadata.

@@

\paragraph{Automatic Service Detection} The tool builds in a basic service detection process that is executed during node registration and at rare periodic intervals after this point.
 It detects the presence of known services such as apache2, nginx or mysql within the currently executing process table, these services are defined in a simple component definition list bundled with the agent software, making it easy to expand in future. It then takes further steps to extract version numbers from installed services and includes this information in the node metadata.

@@

\begin{itemize}
  \item Service - An application that provides functionality to an end-user or another service over a network. eg.\ a website or API.
  \item Component - Part of a service which performs a specific function, eg.\ an application server or a load balancer.
  \item Host - A physical, virtual or otherwise isolated container responsible for running some or all of the components of a service.
  \item Metric - A measure of a characteristic at a given point in time, eg.\ CPU load or average request latency. Typically quantitative in nature.
  \item Check - Validation that an arbitrary property holds at a given point in time, eg.\ A specific service is listening on port 80. Typically qualitative in nature.
  \item Event - A discrete action performed upon the system, eg.\ a code deployment or service re-configuration. Events can be correlated with metrics.
  \item Incident - A specific type of event which is automatically triggered by the failure of one or more checks, incidents correlate recent events with failed checks and can be related to previous incidents of a similar nature to aid in root-cause analysis.
  \item Context - Additional data which aids root-cause analysis, eg.\ log files from a malfunctioning service.
\end{itemize}

@@

The streaming service manages the collection of test results, metrics and other events such as code deployments via event streams over TCP or UDP. Streams reporting errors or anomalous metrics will trigger status changes in the management API, as will streams that stop recieving messages, as this may indicate failure in the monitoring framework itself.

@@

  \item{instance\_eval} - This function executes a given block in the context of the recieving object. This means that lots of typical structural techniques such as defining classes and methods can be hidden as part of the implementation, and relied upon during execution. An example of this is included below:

@@

The most important feature of the DSL is the ability to encode checks as Ruby blocks. A block contains one or more assertions that throw exceptions with descriptive error messages if they fail. Below is a simple example of this. The health check will be executed every 30 seconds by the monitoring agent. The monitoring agent will catch any exceptions and return a descriptive error message which may be printed to the console or serialized and sent to the event stream.

@@

\subsubsection{Installation}

@@

\begin{enumerate}
  \item{Detection of Operating System} - The tool initially supports Linux and Mac OS X only, it will distinguish between these operating systems or abort the installation if running on a different platform.
  \item{Detection or Installation of Ruby} - If Ruby 2.0.1 is not installed, the script will use the standard package manager of the distribution to install it.
  \item{Download of Agent Software} - The source code will be obtained directly from GitHub, dependencies will be resolved and installed.
  \item{Host Registration} - Optionally, the host can be registed with the management server, the hostname and IP address as well as an optional service name and environment will be passed to the management server. This is appropriate on servers but not development workstations.
\end{enumerate}

@@

\begin{description}
  \item[Check] - Run configured health checks on demand. This command is typically used on developer workstations as they iteratively build a service specification using the DSL\. It provides fast feedback similar to running a suite of unit tests and can be used to quicky validate assumptions, before pushing the specification to the management server.
  \item[Discover] - Creates a service specification from detected system services. This feature can be used primarily to reverse-engineer existing infrastructure setups into a service specification. More details about the discovery mechanism are provided in the next subsection.
  \item[Graph] - Render a graph of the current service - This can be primarly used on the developer workstation, it will use the GraphViz library to render a graph of the system and its dependenciesin image or PDF format. This can be used on its own as a form of documentation, but is typically intended to help visualise the service specification and its dependencies to ensure its correctness.
  \item[Push] - Pushes the specified service to a management server - This command will parse the DSL and serialize each object to JSON, it is then sent as a PUT request to the management API\. Changes to the service specification take effect immediately and update any connected users in near real-time.
  \item[Setup] - Register this host with the management server - This is used on the servers running the system under test, it will associate a host with a defined service, listing it in the Web UI and accepting events and check data from it.
  \item[Start] - Starts a monitoring daemon process which runs in the background and executes checks according to their defined intervals. It will report all data back to the management server's event stream endpoint. Any errors occurring during this process will be logged to a file for later analysis.
  \item[Stop] - Stop monitoring the specified service.
  \item[Restart] - Restart monitoring of the specified service.
  \item[Help] - Describes the function of all available commands.
\end{description}

@@

@@

\begin{enumerate}
  \item The main event stream is filtered down to events with the tag `check'.
  \item The `metric' field of the event is forwarded to Graphite, under an appropriate path, e.g. `example-service.nginx.health'. In the case of health checks, the value is simply 0 or 1, signifying failure or success respectively.
  \item If the most recent event has a different `state' field to the previous event of the same type, queue a notify task. This could happen when a check transitions state due to an error, or where another check packet was not recieved within the expected TTL interval.

@@

>>>>

After taking part in the delivery of software in the highly regulated insurance industry, I noted some problems with the reliability of the software delivered. These reliability issues were generally unrelated to the application itself, but predominantly caused by the growth in complexity of the underlying application infrastructure, which was provided and managed within the company's own datacenter as a result of regulatory constraints. As the application grew in complexity over time, the number of dependent services and components needed to keep it running became unmanageable by the company's own IT staff.

@@

The tool will then execute this model as a means of system monitoring. It will verify the correctness of the live system continuously and warn of any divergence. It will also allow the attachment of meta-data and metrics to nodes. This model, meta-data and metrics will then serve as a useful diagnostic tool if the state of the system changes, e.g.\ a loss of connectivity between two nodes. This fine-grained and clear mechanism of communicating failure should reduce investigation and response times when issues arise, and provide the insight which prevents them from recurring.

@@

The problem of increasing software complexity is a well publicised and researched phenomenon. Software which is complex takes longer to develop, and is harder to maintain. As it affects the value proposition to the user so directly, it is often the dominant factor in determining the feasibility of implementing a given system.

@@

Modelling of complex software has been used for a number of years as a tool of communication between developers, project managers and stakeholders. A number of tools can be used to produce clear diagrams of a system, but models provide a level of semantic meaning that allows them to be practically used to derive or verify an implementation. The Unified Modelling Language is considered the de-facto set of graphic notation techniques to create visual models of object-oriented software-intensive systems~\cite{UMLDefinition}.

@@

Studies into increased quality, maintainability and developer productivity when using MDSE techniques can be found commonly in academia but are difficult to find in industry. In a study into Industrial Adoption of Model Driven Development, it was stated that "interviewees emphasised tool immaturity, complexity and lack of usability as major barriers to adoption"~\cite{IndustryMDSE}.

@@

Nagios is in active use by over 250,000 organisations worldwide, including Amazon, AT\&T and JP Morgan Chase~\cite{NagiosUsage}. It supports a number of common monitoring architectures, including both agent based and agent-less (based on the Simple Network Management Protocol) which means it can run efficiently up to a very large number of hosts~\cite{NagiosArchitectures}.

Anecdotal feedback about Nagios typically suggests that it is a mature and flexible solution for basic monitoring of large numbers of hosts, but severely lacks some higher level concepts such as integration tests of system components. It is also considered extremely time consuming and counter-intuitive to set up and manage, which perhaps explains its lack of adoption within smaller businesses/startups.

@@

\begin{itemize}
  \item A Technical Architect at Red Badger Consulting
  \item A System Administrator at HCL
\end{itemize}

@@

\paragraph{Simplicity and Pragmatism} Often the need for extensive application and infrastructure monitoring is not recognised until after the application has been deployed into a production environment. Introducing monitoring at an earlier stage, e.g.. in a staging environment, can highlight issues of reliability and performance before they are experienced in production. Monitoring key domain specific metrics can also aid in reasoning about the runtime behaviour of the system. With these benefits in mind, the tool should be simple enough to introduce early in the development process with minimal effort on the part of the developer, it should provide pragmatic value from the start, and offer the flexibility to continue to deliver value as the system grows in complexity.

@@

When the binary is first executed, it performs node registration by making a PUT request to the monitoring tool’s REST API with the hostname of the server, and supplementary meta-data.

@@

\paragraph{Automatic Service Detection} The tool builds in a basic service detection process that is executed during node registration and at rare periodic intervals after this point.
 It detects the presence of known services such as Apache2, Nginx or MySQL within the currently executing process table, these services are defined in a simple component definition list bundled with the agent software, making it easy to expand in future. It then takes further steps to extract version numbers from installed services and includes this information in the node meta-data.

@@

\begin{itemize}
  \item Service - An application that provides functionality to an end-user or another service over a network\. e.g. a website or API.
  \item Component - Part of a service which performs a specific function, e.g.\ an application server or a load balancer.
  \item Host - A physical, virtual or otherwise isolated container responsible for running some or all of the components of a service.
  \item Metric - A measure of a characteristic at a given point in time, e.g.\ CPU load or average request latency. Typically quantitative in nature.
  \item Check - Validation that an arbitrary property holds at a given point in time, e.g.\ A specific service is listening on port 80. Typically qualitative in nature.
  \item Event - A discrete action performed upon the system, e.g.\ a code deployment or service re-configuration. Events can be correlated with metrics.
  \item Incident - A specific type of event which is automatically triggered by the failure of one or more checks, incidents correlate recent events with failed checks and can be related to previous incidents of a similar nature to aid in root-cause analysis.
  \item Context - Additional data which aids root-cause analysis, e.g.\ log files from a malfunctioning service.
\end{itemize}

@@

The streaming service manages the collection of test results, metrics and other events such as code deployments via event streams over TCP or UDP. Streams reporting errors or anomalous metrics will trigger status changes in the management API, as will streams that stop receiving messages, as this may indicate failure in the monitoring framework itself.

@@

  \item{instance\_eval} - This function executes a given block in the context of the receiving object. This means that lots of typical structural techniques such as defining classes and methods can be hidden as part of the implementation, and relied upon during execution. An example of this is included below:

@@

The most important feature of the DSL is the ability to encode checks as Ruby blocks. A block contains one or more assertions that throw exceptions with descriptive error messages if they fail. Below is a simple example of this. The health check will be executed every 30 seconds by the monitoring agent. The monitoring agent will catch any exceptions and return a descriptive error message which may be printed to the console or serialised and sent to the event stream.

@@

\subsection{Installation}

@@

\begin{enumerate}
  \item{Detection of Operating System} - The tool initially supports Linux and Mac OS X only, it will distinguish between these operating systems or abort the installation if running on a different platform.
  \item{Detection or Installation of Ruby} - If Ruby 2.0.1 is not installed, the script will use the standard package manager of the distribution to install it.
  \item{Download of Agent Software} - The source code will be obtained directly from GitHub, dependencies will be resolved and installed.
  \item{Host Registration} - Optionally, the host can be registered with the management server, the hostname and IP address as well as an optional service name and environment will be passed to the management server. This is appropriate on servers but not development workstations.
\end{enumerate}

@@

\begin{description}
  \item[Check] - Run configured health checks on demand. This command is typically used on developer workstations as they iteratively build a service specification using the DSL\. It provides fast feedback similar to running a suite of unit tests and can be used to quickly validate assumptions, before pushing the specification to the management server.
  \item[Discover] - Creates a service specification from detected system services. This feature can be used primarily to reverse-engineer existing infrastructure setups into a service specification. More details about the discovery mechanism are provided in the next subsection.
  \item[Graph] - Render a graph of the current service - This can be primarily used on the developer workstation, it will use the GraphViz library to render a graph of the system and its dependencies image or PDF format. This can be used on its own as a form of documentation, but is typically intended to help visualise the service specification and its dependencies to ensure its correctness.
  \item[Push] - Pushes the specified service to a management server - This command will parse the DSL and serialise each object to JSON, it is then sent as a PUT request to the management API\. Changes to the service specification take effect immediately and update any connected users in near real-time.
  \item[Setup] - Register this host with the management server - This is used on the servers running the system under test, it will associate a host with a defined service, listing it in the Web UI and accepting events and check data from it.
  \item[Start] - Starts a monitoring daemon process which runs in the background and executes checks according to their defined intervals. It will report all data back to the management server's event stream endpoint. Any errors occurring during this process will be logged to a file for later analysis.
  \item[Stop] - Stop monitoring the specified service.
  \item[Restart] - Restart monitoring of the specified service.
  \item[Help] - Describes the function of all available commands.
\end{description}

@@

\subsection{Event Schema \& Routing}

@@

\begin{enumerate}
  \item The main event stream is filtered down to events with the tag `check'.
  \item The `metric' field of the event is forwarded to Graphite, under an appropriate path, e.g. `example-service.nginx.health'. In the case of health checks, the value is simply 0 or 1, signifying failure or success respectively.
  \item If the most recent event has a different `state' field to the previous event of the same type, queue a notify task. This could happen when a check transitions state due to an error, or where another check packet was not received within the expected TTL interval.

@@

\section{Management API}
\subsection{Ruby on Rails}
\subsubsection{ActiveRecord}
\subsection{Incident Management}
\subsection{Live Updates - SSE \& WebSockets}
\subsection{Email and SMS Notifications}

\section{Web UI}

\subsection{React.js}
\subsection{Data Visualisation - D3.js}
\subsection{Live Updates}
